---
title: "honors_models"
author: "Ava Braiter"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load in Packages

```{r}
library(tidyverse)
library(readxl)
library(mosaic)
library(car)
library(stats)
library(ggplot2)
library(GGally)
library(leaps)

```

#ERA 1

### Load in Data

```{r}
era_1 <- read_xlsx("era1.xlsx")


```

### Make Percent Change Variables

```{r}
pc_1 <- era_1 %>% 
  mutate(vc_pc=(venture_capital_2001-venture_capital_1996)/venture_capital_1996,
         patents_pc=((patents_2001-patents_1996)/patents_1996),
         firm_death_pc=((firm_deaths_2001-firm_deaths_1996)/firm_deaths_1996),
         bach_degrees_pc=(bach_degrees_2001-bach_degrees_1996)/bach_degrees_1996,
         net_jobs_pc=(net_jobs_2001-net_jobs_1996)/net_jobs_1996,
         edu_investment_pc=(edu_investment_2001-edu_investment_1996)/edu_investment_1996,
         info_net_jobs_pc=(info_net_jobs_2001-info_net_jobs_1996)/info_net_jobs_1996,
         employment_pc=(employment_2001-employment_1996)/employment_1996,
         gdp_pc=(gdp_2001-gdp_1996)/gdp_1996,
         gdp_per_capita_pc=(gdp_per_capita_2001-gdp_per_capita_1996)/gdp_per_capita_1996,
         population_pc=(population_2001-population_1996)/population_1996,
         info_employment_pc=(info_employment_2001-info_employment_1996)/info_employment_1996)

pc_1 <- pc_1 %>% 
  select(state, info_employment_2001, vc_pc, patents_pc, firm_death_pc, bach_degrees_pc, net_jobs_pc, edu_investment_pc, info_net_jobs_pc, employment_pc, gdp_pc, gdp_per_capita_pc, population_pc, info_employment_pc)

#make inf and NaN into NA 
pc_1 <- pc_1 %>% 
  mutate_all(~ifelse(is.nan(.), NA, .))
pc_1[sapply(pc_1, is.infinite)] <- NA

```

#Tech Region Models

Selected 25 states with the most information sector employment in 2001 for high df

For high tech regions, Illinois, Virginia, and North Carolina were removed because they were extreme outliers for venture capital percent change.

```{r}

pc_1_high <- pc_1 %>% 
  arrange(desc(info_employment_2001))
pc_1_high <- pc_1_high %>% 
  slice(c(1:4, 6, 8:15, 17:25))

pc_1_low <- pc_1 %>% 
  arrange(desc(info_employment_2001)) %>% 
  slice(26:51)
pc_1_low_no <- pc_1_low %>% 
  slice(c(1:9, 12:15, 17, 19:26))

```

##High Tech

Dependent Variable: information sector employment

###Correlation plots

```{r, warnings=FALSE}
#high tech areas
ggpairs(pc_1_high, columns=c("edu_investment_pc", "vc_pc", "bach_degrees_pc", "gdp_pc", "patents_pc", "firm_death_pc", "info_employment_pc"))

```

###Education Investment

```{r, warning=FALSE}
educ.high.fit.1a <- lm(info_employment_pc~edu_investment_pc, data=pc_1_high, na.rm=TRUE)
summary(educ.high.fit.1a)

educ.high.fit.1b <- lm(info_employment_pc~edu_investment_pc+gdp_pc, data=pc_1_high, na.rm=TRUE)
summary(educ.high.fit.1b)

educ.high.fit.1c <- lm(info_employment_pc~edu_investment_pc+gdp_pc+patents_pc, data=pc_1_high, na.rm=TRUE)
summary(educ.high.fit.1c)

educ.high.fit.1d <- lm(info_employment_pc~edu_investment_pc+bach_degrees_pc, data=pc_1_high, na.rm=TRUE)
summary(educ.high.fit.1d)
```

###Patents

```{r, warning=FALSE}
patents.high.fit.1a <- lm(info_employment_pc~patents_pc, data=pc_1_high, na.rm=TRUE)
summary(patents.high.fit.1a)

patents.high.fit.1b <- lm(info_employment_pc~patents_pc+edu_investment_pc, data=pc_1_high, na.rm=TRUE)
summary(patents.high.fit.1b)

patents.high.fit.1c <- lm(info_employment_pc~patents_pc+edu_investment_pc+bach_degrees_pc, data=pc_1_high, na.rm=TRUE)
summary(patents.high.fit.1c)
```

### Venture Capital

```{r, warning=FALSE}
vc.high.fit.1a <- lm(info_employment_pc~vc_pc, data=pc_1_high, na.rm=TRUE)
summary(vc.high.fit.1a)

vc.high.fit.1b <- lm(info_employment_pc~vc_pc+edu_investment_pc, data=pc_1_high, na.rm=TRUE)
summary(vc.high.fit.1b)

vc.high.fit.1c <- lm(info_employment_pc~vc_pc+edu_investment_pc+gdp_pc, data=pc_1_high, na.rm=TRUE)
summary(vc.high.fit.1c)

vc.high.fit.1d <- lm(info_employment_pc~vc_pc+patents_pc, data=pc_1_high, na.rm=TRUE)
summary(vc.high.fit.1d)

```

###Firm death
```{r,warning=FALSE}
fd.high.fit.1a <- lm(info_employment_pc~firm_death_pc, data=pc_1_high, na.rm=TRUE)
summary(fd.high.fit.1a)

fd.high.fit.1b <- lm(info_employment_pc~firm_death_pc+edu_investment_pc, data=pc_1_high, na.rm=TRUE)
summary(fd.high.fit.1b)

fd.high.fit.1c <- lm(info_employment_pc~firm_death_pc+edu_investment_pc+bach_degrees_pc, data=pc_1_high, na.rm=TRUE)
summary(fd.high.fit.1c)

fd.high.fit.1d <- lm(info_employment_pc~firm_death_pc+vc_pc, data=pc_1_high, na.rm=TRUE)
summary(fd.high.fit.1d)

fd.high.fit.1e <- lm(info_employment_pc~firm_death_pc+patents_pc, data=pc_1_high, na.rm=TRUE)
summary(fd.high.fit.1e)


```

For states with a high amount of information sector employment, patents is the best predictor of information sector employment. Adding additional variables makes the model worse, however, the model is not incredibly strong to begin with (Adj. R-squared = .2747)


```{r}
#look at all models?
best.model.subset1 <- regsubsets(info_employment_pc~vc_pc+patents_pc+firm_death_pc+bach_degrees_pc+edu_investment_pc+gdp_pc, data=pc_1_high, method="exhaustive")
all.models1 <- summary(best.model.subset1)
head(all.models1)


```
other options: patents, firm death, gdp
```{r, warning=FALSE}
best.r2 <- lm(info_employment_pc~edu_investment_pc+firm_death_pc+vc_pc+patents_pc+bach_degrees_pc+gdp_pc, data=pc_1_high, na.rm=TRUE)
summary(best.r2)
```

##Low Tech

```{r, warning=FALSE}
ggpairs(pc_1_low_no, columns=c("edu_investment_pc", "vc_pc", "bach_degrees_pc", "gdp_pc", "patents_pc", "info_employment_pc"))
```
idk if running models are worth it since there's nothign significant here other than gdp

###find best model
```{r, warning=FALSE}
best.low.model.subset1 <- regsubsets(info_employment_pc~vc_pc+patents_pc+firm_death_pc+bach_degrees_pc+edu_investment_pc+gdp_pc, data=pc_1_low, method="exhaustive")
all.low.models1 <- summary(best.low.model.subset1)
head(all.low.models1)
```


#All states

```{r, warning=FALSE}
ggpairs(pc_1, columns=c("vc_pc", "patents_pc", "firm_death_pc", "edu_investment_pc", "gdp_pc", "info_employment_pc"))
```
```{r, warning=FALSE}
pc_1_1 <- pc_1 %>% 
  slice(c(1:8, 10:12, 15:33, 35:39, 41:46, 48:51))
ggpairs(pc_1_1, columns=c("vc_pc", "patents_pc", "firm_death_pc", "edu_investment_pc", "gdp_pc", "population_pc", "info_employment_pc"))
```

###find best of all models

```{r, warning=FALSE}
best.model.subset <- regsubsets(info_employment_pc~vc_pc+patents_pc+firm_death_pc+bach_degrees_pc+edu_investment_pc+population_pc+gdp_pc, data=pc_1_1, method="exhaustive")
all.models <- summary(best.model.subset)
head(all.models)
```

###testing best adj r^2 models
1: gdp_pc (0.365)
2: firm_death_pc+gdp_pc (0.4645)
3: patents_pc+firm_death_pc+gdp_pc (0.4878)
4: patents_pc+firm_death_pc+bach_degrees_pc+gdp_pc (0.4886)
5: patents_pc+firm_death_pc+bach_degrees_pc+edu_investment_pc+gdp_pc (0.4745)
6: vc_pc+patents_pc+firm_death_pc+bach_degrees_pc+edu_investment_pc+gdp_pc (0.4577)
```{r, warning=FALSE}
best.fit.1 <- lm(info_employment_pc~gdp_pc, data=pc_1_1, na.rm=TRUE)
summary(best.fit)

best.fit.2 <- lm(info_employment_pc~firm_death_pc+gdp_pc, data=pc_1_1, na.rm=TRUE)
summary(best.fit.2)

best.fit.3 <- lm(info_employment_pc~patents_pc+firm_death_pc+gdp_pc, data=pc_1_1, na.rm=TRUE)
summary(best.fit.3)

best.fit.4 <- lm(info_employment_pc~patents_pc+firm_death_pc+bach_degrees_pc+gdp_pc, data=pc_1_1, na.rm=TRUE)
summary(best.fit.4)

best.fit.5 <- lm(info_employment_pc~patents_pc+firm_death_pc+bach_degrees_pc+edu_investment_pc+gdp_pc, data=pc_1_1, na.rm=TRUE)
summary(best.fit.5)

best.fit.6 <- lm(info_employment_pc~vc_pc+patents_pc+firm_death_pc+bach_degrees_pc+edu_investment_pc+gdp_pc, data=pc_1_1, na.rm=TRUE)
summary(best.fit.6)

```


firm death with gdp is a very significant predictor with an adjusted r2 of 0.51. adding patents doesn't really affect significance, but adjusted r2 increases to 0.5353.

###ANOVA to test which model is the best
```{r, warning=FALSE}
anova(best.fit.1, best.fit.2)
summary(best.fit.2)
```
The best model for predicting information sector employment growth is model 2: firm_death_pc and gdp_pc. For every one percent increase in firm death pc while holding gdp pc constant, information sector grew by 1.4 percent
